---
title: "p8105_hw2_hy2669"
author: "Haoyang Yi"
date: "2020/9/27"
output: github_document
---
```{r}
library(tidyverse)
library(readxl)
```

## Problem 1
```{r}
trashwheel = read_xlsx('./data/problem1/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
                      sheet = 'Mr. Trash Wheel',
                      range = cell_cols("A:N")) 
trashwheel = janitor::clean_names(trashwheel) %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls)) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  print() # there are 14 observations
# read precipitation of 2017 and 2018 separately
precip_2018 = read_xlsx('./data/problem1/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
                      sheet = '2018 Precipitation',skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = read_xlsx('./data/problem1/Trash-Wheel-Collection-Totals-8-6-19.xlsx',
                      sheet ='2017 Precipitation',skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
# combine annual precipitation
month_df = tibble(month = 1:12,
                month_name = month.name)
precip_df = bind_rows(precip_2018,precip_2017)
left_join(precip_df,month_df,by ='month') 

sum(precip_df$total[precip_df$year == 2018])
median(trashwheel$sports_balls[trashwheel$year == 2017])
```
  There are 4 observations in precipitation dataset for 2017 and 2018.There are 14 observations in Mr.Trash Wheel sheet dataset. The total precipitation in 2018 is `r sum(precip_df$total[precip_df$year==2018])`. The median number of sports balls in a dumpster in 2017 is `r median(trashwheel$sports_balls[trashwheel$year==2017])`.

## Problem 2
```{r}
nyc_transit_df = read_csv('./data/problem2/NYC_Transit_Subway_Entrance_And_Exit_Data.csv')  
  nyc_transit_df = janitor::clean_names(nyc_transit_df) %>% 
  select(line,station_name,station_latitude,station_longitude,
         route1:route11,entry,vending,entrance_type, ada) %>%
  mutate(entry = recode(entry,"YES"="TRUE","NO"="FALSE")) %>%
  mutate(entry = as.logical(entry)) 
  str(nyc_transit_df)
``` 
   This dataset contains information of stations including names,locations,routes they serve,entrance type, ADA compliance,entry and vending status. 
  Data cleaning steps: read the data-clean the name-select columns to retain-convert the entry variable from character to logical.The resulting dataset has `r ncol(nyc_transit_df)` columns and `r nrow(nyc_transit_df)` rows.These data are not tidy since there are too many route columns, and route1:route7 are character variables while route8:route11 are numeric variables.
  
```{r}   
  nyc_station<-distinct(nyc_transit_df,station_name,line) # get distinct stations based on name and line.
  nrow(nyc_station)# there are 465 distinct stations.
  count(nyc_transit_df,ada) # 468 stations are ADA compliant.
  nyc = filter(nyc_transit_df,vending =='NO') # station without vending 
  nycentry = filter(nyc,entry == TRUE) # station without vending allow entrance
  nrow(nycentry)/nrow(nyc) # the proportion = number of stations without vending allow entrance/ total number of stations without vending = 0.377
  nyc_tidy_df = nyc_transit_df %>%
  mutate_at(vars(route1:route11),as.character) %>%
  pivot_longer(route1:route11,names_to ='route_name',names_prefix ='route',
               values_to ='route_number') %>%
  drop_na(route_number) %>%
  print() # Reformat the data so that the route number and route name are distinct variables
  a_df<-filter(nyc_tidy_df,route_number =='A')
  nrow(distinct(a_df,station_name,line)) # there are 60 distinct stations serve the A train
  ada_df = filter(a_df,ada =='TRUE')
  nrow(distinct(ada_df,station_name,line)) # Of the stations that serve the A train,17 are ADA compliant
```

## Problem 3
```{r}
pols_month_df = read_csv('./data/problem3/pols-month.csv') 
snp_df = read_csv('./data/problem3/snp.csv')
unemployment_df = read_csv('./data/problem3/unemployment.csv')
```
```{r}
  pols_month_df = separate(pols_month_df,mon,into = c('year','month','day')) %>%
  mutate(month = month.abb[as.factor(month)]) %>% # replace month number with month name
  mutate(president = recode(prez_gop,'0'='dem','1'='gop','2'='abnormal')) %>% 
  mutate(year = as.integer(year)) %>%  # convert year variable to integer 
  select(-prez_dem,-prez_gop,-day) # separate mon into year month and day. 
# Create president variable to show the party that the president belongs. remove prez_dem, prez_gop and day variables
  snp_df = separate(snp_df,date,into=c('month','day','year')) %>%
  arrange(year,month) %>% # arrange according to year and month
  mutate(month = month.abb[as.factor(month)]) %>% 
  mutate(year = as.integer(year)) %>% 
  select(year,month,close)  # clean and tidy the data in snp.csv,organize columns.
  unemployment_df = pivot_longer(unemployment_df,Jan:Dec,names_to = 'month',
                                 values_to = 'rate of unemployment') %>%
  rename(year = Year) # switch the unemployment dataset to 'long'
  # rename the Year variable to year according to name of variable in pols_month and snp datasets.
two_df = left_join(pols_month_df,snp_df) 
final_df = left_join(two_df,unemployment_df) # first, merge snp into pols, then merge unemployment in to the previous result.
```
  The pols dataset contains numbers of republican and democratic national politicians on associated date.The snp dataset contains closing values fo S&P stock index on the associated date. The unemployment dataset contains percentage of unemployment in each month of the associated year.
  `r skimr::skim(final_df)`
  The dimension of the resulting dataset is 822 rows and 11 columns. The range of years is 1947-2015. Key variables are president(republican or democratic), numbers of republican and democratic governors,senators and representatives, closing values of S&P stock index, precentage of unemployment.


